{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-10-02 thesis/future.tex
future test case design

Each of the previous three chapters concluded by
discussing specific limitations,
listing concrete and immediate ways to address them
with existing techniques
(\sect{\ref{sec:quicksand-discussion}},
\sect{\ref{sec:education-discussion}},
\sect{\ref{sec:tm-discussion}}).
Meanwhile, this chapter takes a broader interpretation of ``future work'',
namely,
how might future Landslides solve research problems I didn't even pose to begin with.
%The reader looking for their own thesis topic would most likely find it here.

@@

@@

% envision a framework that uses some code coverage tester to guide choice of pps and/or the user's without_function calls 
% to get that coverage -- although relating flow control coverage directly to concurrency coverage is not straightforward
% framework can also use 'git diff's, like comparing the most recent commit against an already-trusted codebase,
% to figure out how to isolate just the stuff the change affects (memory intersection & data flow resulting from those conflix)
% and then automatically assign some without_functions/etc to test just that commit

% failure injection, testing malloc/etc failures
% eg student condvars which loop around malloc failures can deadlock
% can use retry set reduction in principle to prune the same kind of equivalences

% also here we've got discussion of the exit-UAF buggers where landslide oops verifs the test anyway
% because they're using a free pool and not free()
% need a 3rd thread at least, to expose this (by actually getting corruption)

>>>>

Each of the previous three chapters concluded by
discussing specific limitations,
listing concrete and immediate ways to address them
with existing techniques
(\sect{\ref{sec:quicksand-discussion}},
\sect{\ref{sec:education-discussion}},
\sect{\ref{sec:tm-discussion}}).
Meanwhile, this chapter takes a broader interpretation of ``future work'',
namely,
how might future Landslides solve research problems I didn't even pose to begin with.
%The reader looking for their own thesis topic would most likely find it here.
%The reader should beware of wild speculation ahead
%which may assume solutions to extremely difficult problems with a wave of the hand.

@@

In its current form, Landslide is limited to testing only those behaviours that the test case it's hooked up to can generate.
The most obvious limitation of this is resource exhaustion scenarios:
stateless model checking simply cannot handle tests long-running enough to exhaust system memory
(succumbing to exponential explosion long before),
% "involve" here, for subjunctive?
so cannot exercise any flow control that involves {\tt malloc()} failing, for example.
This specific issue could be solved using by-hand annotations to block all preemption points until just before exhaustion,
%(requires model-checking expertise on the programmer's part)
or by extending Landslide to inject allocation failures at {\tt malloc()} callsites (akin to {\tt \_xbegin()}).
%(assumes certain behavioural properties of the allocation library).
However,
these require the user to realize in advance that she should worry about allocation-failure bugs,
and to configure Landslide specifically to target them.
In general, a mature testing tool should not require the user to
know in advance where her bugs might be
before being able to conduct an effective test.

In future work, a stateless model checking framework could collaborate with its user to semi-automatically design better tests.
% i hate the way maple actually does it, in principle, but i need to cite smth here
Concurrency coverage metrics such as proposed by Maple \cite{maple}
% wild speculation
could be extended to capture possible program behaviours under any test input,
not just within the fixed state space under one given test,
and symbolic execution frameworks such as Contessa \cite{contessa}
could search for possible test inputs to suggest changes to the test case which might expose them.
In cases where not-yet-covered conditional branches require
certain function return values,
such as {\tt malloc()} failing,
the tool could offer to add failure injection with the user's approval,
or at least prompt her to write a new test with that as a subgoal.

In the case study of submitted P2 bugs (\sect{\ref{sec:education-eval-bug-case-study}}),
I noticed several submissions of a common pattern
in which Landslide overlooked the {\tt thr\_exit()} use-after-free bug:
they all manually recycled the exited thread's stack using an internal free-list,
rather than calling {\tt free()} or {\tt remove\_pages},
so Landslide's heap-checker failed to see anything out of the ordinary about the subsequent writes to it.
Explicit annotations about the custom free-list's invariants could make Landslide treat it like {\tt free()}
and catch the errant write,
but that begs the question of knowing about the bug in advance.
Catching this bug would require extending the test case to {\tt thr\_fork()} a new thread
to reuse the old stack and suffer data corruption from the stray write.
In this case, the user might happen to find this just by increasing the number of threads/iterations,
which is just good verification principles,
but in general, automatically inferring such data structure invariants
to suggest better assertions
% is this true??
remains an open problem.

@@
